---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all(".", quiet = TRUE)
```

# scenario2risk

<!-- badges: start -->
<!-- badges: end -->

`scenario2risk` estimates downside portfolio risk from simulated
macro-financial scenarios. The package combines principal component analysis
with a factor-augmented VAR model to generate future equity, bond, and cash
return paths, then summarizes terminal-loss risk and rolling VaR exceedance
diagnostics.

It is designed around two main user-facing functions:

- `portfolio_risk()` for scenario-based portfolio loss metrics
- `model_check()` for rolling VaR exceedance checks against a historical
  simulation benchmark

## Installation

You can install the development version from GitHub once the repository is
online:

``` r
# install.packages("remotes")
remotes::install_github("YOUR_GITHUB_USERNAME/scenario2risk")
```

Replace `YOUR_GITHUB_USERNAME` with your GitHub account name after you publish
the repository.

If you are working from a local copy of the project, you can also install it
from source:

``` r
# install.packages("remotes")
remotes::install_local(".")
```

## Example: Portfolio Risk

The example below estimates terminal-loss risk for a simple three-asset
portfolio using the package's built-in demo data.

```{r example-risk, fig.height=4.5, fig.width=7, fig.cap="Simulated terminal portfolio loss distribution."}
risk <- portfolio_risk(
  macro_data = demo_macro_data,
  return_data = demo_return_data,
  portfolio = demo_portfolio_weights,
  horizon = 12,
  n_scenarios = 200,
  k = 2,
  var_lag = 1,
  seed = 123
)

as.data.frame(risk$risk_summary)
plot_portfolio_risk(risk)
```

## Example: Rolling Model Check

`model_check()` compares scenario-based VaR exceedance rates with a historical
simulation benchmark across rolling forecast origins.

```{r example-check}
check <- model_check(
  macro_data = demo_macro_data,
  return_data = demo_return_data,
  portfolio = demo_portfolio_weights,
  horizon = 6,
  n_scenarios = 50,
  k = 2,
  var_lag = 1,
  seed = 123,
  window = 120,
  n_origins = 6
)

as.data.frame(check$var_exceedance)
```

## Included Example Data

The package includes three small datasets for reproducible examples and tests:

- `demo_macro_data`: monthly macro predictor series
- `demo_return_data`: monthly equity, bond, and cash returns
- `demo_portfolio_weights`: example portfolio weights for the three assets

## Development Notes

`README.md` is generated from `README.Rmd`. After editing the source file, run
the following to refresh the GitHub-facing markdown file:

``` r
devtools::build_readme()
```
